WDIAS : A Microservices-Based Weather Data Integration and Assimilation System\\

### 2020-05-24 lit/sec_1-fews.tex
Proofread chapter 2

Deltares developed \acrshort{fews} \cite{Werner2013TheSystem} in Netherland for operational forecasting agencies. The \acrshort{fews} codebase is not fully open source at the moment. The main objective of a forecasting system is to run multiple NWMs, hydrologic, and hydrodynamic models which are required for the predictions. This approach is known as a model-centric approach \cite{Werner2005FloodCatchments}. In the model-centric approach, the inputs need to be in the format specific to the model. Also, the outputs produced by the models are specific to the model; hence, difficult to integrate with other models. For example, FLO2D creates human-readable text output files. However, the Delft-FEWS uses a more modular approach, which is easier to integrate new models and redevelop the system for new requirements. Thus, Delft-FEWS is an integration framework or middleware for the models, rather specific model workflow for weather forecasting.

@@

Are proposed by Haggett \cite{Haggett1998AnWales} key elements of a forecasting system are defined as detection, forecasting, dissemination and warning, and responding to the events. The \acrshort{fews} focuses on the forecasting step out from above steps. The main objective of this step is to provide additional lead time by forecasting future hydro-meteorological conditions \cite{Werner2005FloodCatchments}. It is a valid argument that providing accurate predictions with greater lead time can reduce the level of destruction. To forecast, the system should be capable of integrating real-time data from meteorological and hydrological observation network systems, and the disseminate the forecasted results through relevant products to the warning process.

@@

\cref{fi:fews_schematic} shows the connection between a forecasting system to real-time data integration systems and dissemination systems. One of the widely practiced use cases is to use meteorological forecast data to estimate precipitation, and then use a hydrological and hydraulic model chain to predict the surface water level affected. The hydrological and hydrodynamic models should design based on geographical data. The ground surface should be examined geographically and divided into catchments. Based on the affected catchment, it can be further divided into sub-catchments to reduce the complexity of the simulation task, and reduce simulation time with parallel processing. Ideally, the forecasting system flexible enough to allow modification of models and data while keeping the system interaction method constant as working with multiple forecasters.

@@

The \acrshort{fews} follows a data-centric approach, where it provides a common data-model to interact with other components. All the timeseries data stored in a database by converting them to a common data model. New models are integrated into the system by using one of the interfaces provided to interact with the common data-model \cite{Werner2013TheSystem}. The common data-model allows storing data efficiently without duplicates. Then, the system supports operations like creating forecast reports and sharing the data by integrating new codebase as adapters. However, this approach cannot handle multiple data formats. The \acrshort{fews} overcome this issue by introducing adapters for most of the commonly used data formats such as CSV, NetCDF, and JSON. While this enables users to import data into the system; it also introduces adding additional complexity to the system.

The timeseries data is one of scalar, vector, or grid data type. And different types of data stored as binary objects in a timeseries table inside the \acrshort{fews} system. Functional components or any integrated models do not have direct access to the timeseries table, and those should use the data access module to access the data \cite{Werner2013TheSystem}. As explained in the system interpretation, all data access requests need to go through the data access module cause to increase the cost of data access. Because the \acrshort{fews} stores different data types as binary objects, there is a penalty for converting data into binary objects and vice versa. Since all timeseries data is stored within a single data source, it allows the system to handle data without much system complexity. However, a single data source becomes a bottleneck when it is required to fulfill all the timeseries data requirements. Further, the performance goes down due to the need to scream Spatio-temporal weather data.

Given the above drawback on storing the data in the system, the Delft-FEWS data model uses a set of primary attributes from weather data to identify a timeseries uniquely. Those primary attributes are location, data type, and source of the data \cite{Werner2013TheSystem}. The primary attributes are always required, and there are some secondary attributes to add timeseries metadata further. Also, primary attributes used to index data for making data queries faster. Moreover, Delft-FEWS is able to create separate indexes for each attribute while storing large amounts of data. As an example, the system can separate timeseries data into multiple databases based on the source of timeseries, such as an external source of data or a hydrological model, which creates the timeseries and then index each database on other attributes. Alternatively, separate timeseries by data type and stored in multiple storages give the capability for the system to scale with a factor of identical types. An example of separation by data types of scalar, vector, and gridded will increase the scaling factor by 3x.

Data preprocessing and manipulation are necessary processes in weather forecasting. The data integrated from external sources are not in the appropriate temporal and spatial format that can directly feed as an input to forecasting models or use in other applications. Therefore, the forecasting environment provides generic data processing steps to convert data into model compatible formats. Such examples are serial and spatial interpolation, data validation, aggregation and disaggregation, and data fusion \cite{Werner2013TheSystem}. This is a vital feature in a forecasting system and affects the quality and accuracy of the predicted data outputs. Because of the common data model concept in the Delft-FEWS, data processing via these functions are much useful. The system contains a set of default data processing functions, but required algorithms can implement as a new Java class integrated with the \acrfull{api} provided by the Delft-FEWS. The Delft-FEWS has language dependency while additional features integration since users need to implement the new functional extensions using Java.

>>>>

Deltares developed \acrshort{fews} \cite{Werner2013TheSystem} in Netherland for operational forecasting agencies. The \acrshort{fews} codebase is not fully open source at the moment. The main objective of a forecasting system is to run multiple NWMs, hydrologic models, and hydrodynamic models which are required for the predictions. This approach is known as a model-centric approach \cite{Werner2005FloodCatchments}. In the model-centric approach, the inputs need to be in the format specific to the model. Also, the outputs produced by the models are specific to the model; hence, difficult to integrate with other models. For example, FLO2D creates human-readable text files as output. However, the Delft-FEWS uses a more modular approach, which is easier to integrate new models and update the system for new requirements. Thus, Delft-FEWS is an integration framework or middleware for the models, rather specific model workflow for weather forecasting.

@@

@@

\cref{fi:fews_schematic} shows the connection between a forecasting system to real-time data integration systems and dissemination systems. One of the widely practiced use cases is to use meteorological forecast data to estimate precipitation, and then use a hydrological and hydraulic model chain to predict the surface water level affected. The hydrological and hydrodynamic models should design based on geographical surface data. The ground surface should be examined geographically and divided into catchments. Based on the affected catchment, it can be further divided into sub-catchments to reduce the complexity of the simulation task, and reduce simulation time with parallel processing. Ideally, the forecasting system flexible enough to allow modification of models and data while keeping the system interaction method constant as working with multiple forecasters.

@@

The \acrshort{fews} follows a data-centric approach, where it provides a common data-model to interact with other components. All the timeseries data stored in a database by converting them to the common data model. New models are integrated into the system by using one of the interfaces provided to interact with the common data-model \cite{Werner2013TheSystem}. The common data-model allows storing data efficiently without duplicates. Then, the system supports operations like creating forecast reports and sharing the data by integrating new codebase as adapters. However, this approach cannot handle multiple data formats. The \acrshort{fews} overcome this issue by introducing adapters for most of the commonly used data formats such as CSV, NetCDF, and JSON. While this enables users to import multiple data formats into the system; it also introduces adding additional complexity to the system.

The timeseries data is one of scalar, vector, or grid data type. And different types of data stored as binary objects in a timeseries table inside the \acrshort{fews} system. Functional components or any integrated models do not have direct access to the timeseries table, and those should use the data access module to access the data \cite{Werner2013TheSystem}. As explained in the system interpretation, all data access requests need to go through the data access module cause to increase the cost of data access. Because the \acrshort{fews} stores different data types as binary objects, there is a penalty for converting data into binary objects and vice versa. Since all timeseries data is stored within a single data source, it allows the system to handle data without much system complexity. However, a single data source becomes a bottleneck when it is required to fulfill all the timeseries data requirements. Further, the performance goes down due to the need to stream spatio-temporal weather data.

Given the above drawback on storing the data in the system, the Delft-FEWS data model uses a set of primary attributes from weather data to identify a timeseries uniquely. Those primary attributes are location, data type, and source of the data \cite{Werner2013TheSystem}. The primary attributes are always required, and there are some secondary attributes to add timeseries metadata further. Also, primary attributes used to index data for making data queries faster. Moreover, Delft-FEWS is able to create separate indexes for each attribute while storing large amounts of data. As an example, the system can separate timeseries data into multiple databases based on the source of timeseries, such as an external source of data or a hydrological model, which creates the timeseries and then index each database on other attributes. Alternatively, separate timeseries by data type and stored in multiple storages give the capability for the system to scale with a factor of identical types. An example of separation of database by data types of scalar, vector, and gridded into three independently hosted databases will increase the performance factor by three times.

Data preprocessing and manipulation are necessary processes in weather forecasting. The data integrated from external sources are not in the appropriate temporal and spatial format that can directly feed as an input to forecasting models or use in other applications. Therefore, the forecasting environment provides generic data processing steps to convert data into model compatible formats. Such examples are serial and spatial interpolation, data validation, aggregation and disaggregation, and data fusion \cite{Werner2013TheSystem}. This is a vital feature in a forecasting system and affects the quality and accuracy of the predicted data outputs. Because of the common data model concept in the Delft-FEWS, data processing via these functions are much useful. The system contains a set of default data processing functions, but required algorithms can implement as a new Java class integrated with the \acrfull{api} provided by the Delft-FEWS. The Delft-FEWS has language dependency while additional features integration since users need to implement the new functional extensions using Java programming language.

### 2020-05-24 lit/sec_2-lead.tex
Proofread chapter 2

\acrfull{lead} \cite{Droegemeier2005Service-OrientedWeather} addresses the fundamental research challenges needed to create an integrated and scalable framework for adaptive analysis and prediction of the atmosphere. The weather forecasting researchers require many computer resources to predict and analyze weather models in the mesoscale with less time. Rather than each researcher run and handle their computer resources for the weather experiments, LEAD provides a shared pool of resources. Then the researchers can use the resource pool to run their experiment in shorter amounts of time and a more massive scale. At the time, some researchers are developing their experiment forecasting flows, they are not using the resources, and other researchers can run their forecasting flows using available resources. The foundation of the LEAD system is to create a dynamic workflow orchestration and data management in a Web services framework \cite{Droegemeier2005Service-OrientedWeather}.

@@

The second layer contains fundamental tools such as the following to help to link services together \cite{Droegemeier2005Service-OrientedWeather}:
\begin{itemize}
    \item The users interact with LEAD via a Web portal
    \item Data quality control and assimilation happens via ARPS Data Assimilation System (ADAS)
    \item myLEAD is the metadata catalog
    \item Weather Research and Forecast \acrfull{wrf} \cite{MesoscaleMicroscaleMeteorologyLaboratoryWeatherModel}
    \item Algorithm Development and Mining is a set of tools for analyzing observational data, assimilated data sets, and model output
    \item Integrated Data Viewer 
\end{itemize}

@@

As shown in \cref{fi:lead_framework}, LEAD consists of the following sub-components, as well as provides a distributed system for integrating and testing LEAD's components:
\begin{itemize}
    \item User subsystem – comprises the LEAD portal and enable user can access services
    \item Data subsystem – manages data and metadata, all output of digital models produced by operational or experimental models, and information generated by users.
    \item Tools subsystem – provides all IT tools and meteorological tools
    \item Orchestration subsystem – offers the technologies with which users can manage data flows and model execution flows and can create and own output. It also provides links to other software and processes for continuous or on-demand applications.
\end{itemize}

@@

As shown in \cref{fi:lead_soa}, \acrshort{lead} \acrshort{soa} has five different but highly interconnected layers. The bottom layer represents the calculation, application, and raw data sources that are distributed in the LEAD grid and elsewhere. The next level supports web services that provide access to raw services \cite{Droegemeier2005Service-OrientedWeather}. These two layers are working together since LEAD system resources are distributed over multiple locations and creating a pool of resources. The upper layer of the raw resources is abstracting the complexity of managing and accessing the resources and provides simplified access. In the Upper layers, the underline system is viewed as an unlimited resource pool for storing and handling data.

>>>>

\acrfull{lead} \cite{Droegemeier2005Service-OrientedWeather} addresses the fundamental research challenges needed to create an integrated and scalable framework for adaptive analysis and prediction of the atmosphere. The weather forecasting researchers require many computer resources to predict and analyze weather models in the mesoscale with less time. Rather than each researcher run and handle their computer resources for the weather experiments, LEAD provides a shared pool of resources. Then the researchers can use the resource pool to run their experiment in shorter amounts of time and a more massive scale. At the time, some researchers are developing their experiment forecasting flows, they are not using the resources, and other researchers can run their forecasting flows using available resources. The foundation of the LEAD system is to create a dynamic workflow orchestration and data management in a web services framework \cite{Droegemeier2005Service-OrientedWeather}.

@@

The second layer contains fundamental tools such as the following to help to link services together \cite{Droegemeier2005Service-OrientedWeather}:
\begin{itemize}
    \item The users interact with LEAD via a web portal
    \item Data quality control and assimilation happens via ARPS Data Assimilation System (ADAS)
    \item myLEAD is the metadata catalog
    \item \acrfull{wrf} \cite{MesoscaleMicroscaleMeteorologyLaboratoryWeatherModel}
    \item Algorithm development and mining is a set of tools for analyzing observational data, assimilated data sets, and model output
    \item Integrated data viewer 
\end{itemize}

@@

As shown in \cref{fi:lead_framework}, LEAD consists of the following sub-components, as well as provides a distributed system for integrating and testing LEAD's components:
\begin{itemize}
    \item User subsystem – comprises the LEAD portal and enable user can access services
    \item Data subsystem – manages data and metadata, all output of digital models produced by operational or experimental models, and information generated by users
    \item Tools subsystem – provides all IT tools and meteorological tools
    \item Orchestration subsystem – offers the technologies with which users can manage data flows and model execution flows and can create and own output
\end{itemize}

@@

As shown in \cref{fi:lead_soa}, \acrshort{lead} \acrshort{soa} has five different but highly interconnected layers. The bottom layer represents the calculation, application, and raw data sources that are distributed in the LEAD grid and elsewhere. The next level supports web services that provide access to raw services \cite{Droegemeier2005Service-OrientedWeather}. These two layers are working together since LEAD system resources are distributed over multiple locations and creating a pool of resources. The upper layer of the raw resources is abstracting the complexity of managing and accessing the resources and provides simplified access. In the upper layers, the underline system is viewed as an unlimited resource pool for storing and handling data.

### 2020-05-24 lit/sec_3-other-systems.tex
Proofread chapter 2

The data storing mechanism is similar to the other systems discussed above. The DIAS provides an API to import the data and store them in a large array of storage after converting to the required data format. The APIs work to convert or reformat data into manageable forms and are useful for creating the DIAS storage archive. The DIAS's API contains various tools, including the real-time data exploration and archiving system, the data download system, the quality control system, and the metadata management system \cite{Kawasaki2018DataReduction}. To store on the large volume of disk space, it converts the data via a provided set of APIs. The above concept is similar to Delft-FEWs open data model approach. Further, the DIAS contains additional tools that provide the functionality of quality control of the data and metadata management, which is similar to some of the services in LEAD.

As shown in \cref{fi:dias_common_platform}, it offers an application layer that allows scientists to work on research tasks and to develop specific tools and applications by collaborating with IT experts \cite{Kawasaki2018DataReduction}. Once the data is stored on the DIAS system, it shares those data via the Shared API layer and allows users to remote access to the data and easy integration via the APIs.

@@

Based on the above statements, we can conclude that these kinds of functionalities are available at the modern Cloud Computing tools, and out of the box, those tools are providing the scalability and high availability. Using those tools, it is possible to implement such a system with less effort and high confidence.

The MADIS regularly acquires Mesonet data from dozens of network providers representing more than 14,000 stations, and the system translates into more than 30,000 station reports every hour. Different systems that work inside and outside the firewall collect the data. The collected data is sent to the data server of the Advanced Weather-Interactive Processing System (AWIPS) of the central installation for processing and conversion to a common data format, NetCDF. The NetCDF files are then transferred to MADIS compute nodes \cite{Macdermaid2005ArchitectureP2.39}. After reading the statistics of the MADIS system, I realized the paper is a useful reference for system architecture design and its performance testing. When compared to the Delft-FEWS, the MADIS also converts the assimilated data into netCDF format. After converting to netCDF format, the netCDF files are transferred to the compute nodes and accessible via OPeNDAP. Many of these computers are configured using Linux clustering with high availability \cite{Macdermaid2005ArchitectureP2.39}.

>>>>

The data storing mechanism is similar to the other systems discussed above. The DIAS provides an API to import the data and store them in a large array of storage after converting to the required data format. The APIs work to convert or reformat data into manageable forms and are useful for creating the DIAS storage archive. The DIAS's API contains various tools, including the real-time data exploration and archiving system, the data download system, the quality control system, and the metadata management system \cite{Kawasaki2018DataReduction}. To store on the large volume of disk space, it converts the data via a provided set of APIs. The above concept is similar to Delft-FEW's open data model approach. Further, the DIAS contains additional tools that provide the functionality of quality control of the data and metadata management, which is similar to some of the services in LEAD.

As shown in \cref{fi:dias_common_platform}, it offers an application layer that allows scientists to work on research tasks and to develop specific tools and applications by collaborating with IT experts \cite{Kawasaki2018DataReduction}. Once the data is stored on the DIAS system, it shares those data via the shared API layer and allows users to remote access to the data and easy integration via the APIs.

@@

Based on the above statements, we can conclude that these kinds of functionalities are available at the modern cloud computing tools, and out of the box, those tools are providing the scalability and high availability. Using those tools, it is possible to implement such a system with less effort and high confidence.

The MADIS regularly acquires mesonet data from dozens of network providers representing more than 14,000 stations, and the system translates into more than 30,000 station reports every hour. Different systems that work inside and outside the firewall collect the data. The collected data is sent to the data server of the Advanced Weather-Interactive Processing System (AWIPS) of the central installation for processing and conversion to a common data format, NetCDF. The NetCDF files are then transferred to MADIS compute nodes \cite{Macdermaid2005ArchitectureP2.39}. After reading the statistics of the MADIS system, I realized the paper is a useful reference for system architecture design and its performance testing. When compared to the Delft-FEWS, the MADIS also converts the assimilated data into netCDF format. After converting to netCDF format, the netCDF files are transferred to the compute nodes and accessible via OPeNDAP. Many of these computers are configured using Linux clustering with high availability \cite{Macdermaid2005ArchitectureP2.39}.
